---
title: "R Notebook"
output: html_notebook
---

```{r}
PPI_TF_target_Network <- read.csv("Results/Tables/Combined_Networks.tsv", sep = "\t")
```

# Creation of network from functional protein-protein interaction & TF target networks

```{r, include=F}
library(igraph)
library(dplyr)
igraph_PPI_TF_target_Network <- graph_from_data_frame(PPI_TF_target_Network, directed = T)
```

# Eigenvector centrality Definition

In graph theory, eigenvector centrality (also called eigencentrality or prestige score[1]) is a measure of the influence of a node in a network. Relative scores are assigned to all nodes in the network based on the concept that connections to high-scoring nodes contribute more to the score of the node in question than equal connections to low-scoring nodes. A high eigenvector score means that a node is connected to many nodes who themselves have high scores.[2] [3]

Google's PageRank and the Katz centrality are variants of the eigenvector centrality.

For a given graph G:=(V,E) with |V| vertices let A = (a_{v,t}) be the adjacency matrix, i.e. a_{v,t} = 1 if vertex v is linked to vertex t, and a_{v,t} = 0 otherwise. The relative centrality, x, score of vertex v can be defined as:

In graph theory and computer science, an adjacency matrix is a square matrix used to represent a finite graph. The elements of the matrix indicate whether pairs of vertices are adjacent or not in the graph.

# Functions

```{r, include = F}
library(data.table)
library(lucy)

Prepare_features <- function(feature_data_frame, column_of_interest, type_of_data){
  if(type_of_data == "DEG"){
    res <- feature_data_frame[,column_of_interest]
    colnames(res)[1:3] <- c("Gene", "logFC", "P.Value")
  }else{
    res <- feature_data_frame[,column_of_interest]
    colnames(res)[1:3] <- c("Gene", "nes", "pval")
  }
  return(res)
}

Find_most_importants_genes <- function(network){
  res <- list()
  
  ranked_eigen_gene <- network$features[order(-network$features$Eigen_centrality),] %>% head(15) %>% .$Gene
  ranked_page_rank_gene <- network$features[order(-network$features$Page_rank),] %>% head(15) %>% .$Gene
  
  res$ranked_eigen_gene <- ranked_eigen_gene
  res$ranked_page_rank_gene <- ranked_page_rank_gene
  
  V_of_interest <- V(network$network) %>% .[which(names(.) %in% intersect(ranked_eigen_gene, ranked_page_rank_gene))]
  E_of_interest <- E(network$network)[from(V_of_interest) | to(V_of_interest)]

  filtered_graph <- subgraph.edges(network$network, E_of_interest)

  res$network <- filtered_graph
  
  return(res)
}

Prepare_Cytoscape_network <- function(Big_Network = igraph_PPI_TF_target_Network, DEG_analysis, TF_analysis, logFC_treshold = 0.75){
  DEG_of_interest <- DEG_analysis %>% dplyr::filter(abs(logFC) > logFC_treshold & P.Value < 0.1) %>% .$Gene
  TF_of_interest <- TF_analysis %>% dplyr::filter(pval < 0.1) %>% .$Gene

  V_of_interest <- V(Big_Network) %>% .[which(names(.) %in% unique(c(DEG_of_interest, TF_of_interest)))] 

  filtered_graph <- induced_subgraph(Big_Network, V_of_interest)
  

  eigen_centrality_result <- eigen_centrality(filtered_graph, directed = F)$vector

  page_rank_result <- igraph::page.rank(filtered_graph, directed = F)$vector

  features <- merge(DEG_analysis, TF_analysis, by = "Gene", all = T)
  features <- merge(features, eigen_centrality_result, by.x = "Gene", by.y = 0, all = T)
  colnames(features)[ncol(features)] <- "Eigen_centrality"
  features <- merge(features, page_rank_result, by.x = "Gene", by.y = 0, all = T)
  colnames(features)[ncol(features)] <- "Page_rank"

  set(features,which(is.na(features[["nes"]])),"nes",0)
  set(features,which(is.na(features[["pval"]])),"pval",1)
  set(features,which(is.na(features[["Eigen_centrality"]])),"Eigen_centrality",0)
  set(features,which(is.na(features[["Page_rank"]])),"Page_rank",0)

  features$TF <- ifelse(features$nes == 0, F, T)
  
  clustering_eigen <- cluster_leading_eigen(filtered_graph) %>% membership() %>% print() %>% data.frame()
  
  features <- merge(features, clustering_eigen, by.x = "Gene", by.y = 0, all = T)
  set(features,which(is.na(features[["."]])),".",999)
  colnames(features)[ncol(features)] <- "Cluster"
  
  res <- list("features" = features,
              "network" = filtered_graph
              )
  res$most_important_network <- Find_most_importants_genes(res)
  return(res)
}

All_workflow <- function(feature_DEG_df, column_DEG, feature_tf_df, column_TF, NET = igraph_PPI_TF_target_Network, logFC_treshold = 0.75){
  DEG <- Prepare_features(feature_DEG_df, column_DEG, "DEG")
  TF <- Prepare_features(feature_tf_df, column_TF, "TF")
  res <- Prepare_Cytoscape_network(NET, DEG, TF, logFC_treshold)
  return(res)
}
```

# Koichi Data

```{r, include = F}
DF_DEG_Koichi <- read.csv("Results/Tables/NR_R_All_analysis.csv")
DF_TF_Koichi <- read.csv("Results/Tables/Double_MS_Viper_NR_R_REL.csv")




########## NR R

Features_NR_R <- All_workflow(DF_DEG_Koichi, c(1,5,6,2:4), DF_TF_Koichi, c(1:3))

write.table(Features_NR_R$features, "Results/Tables/Genes_of_interest_NR_R.tsv", sep = "\t", row.names = F, quote = F)

Features_NR_R$network %>% 
  igraph::as_data_frame() %>% 
  write.table("Results/Tables/filtered_PPI_TF_NR_R_network.tsv", sep = "\t", row.names = F, quote = F)

Features_NR_R$most_important_network$network %>% 
  igraph::as_data_frame() %>% 
  write.table("Results/Tables/most_important_NR_R_network.tsv", sep = "\t", row.names = F, quote = F)

############ REL R

Features_REL_R <- All_workflow(DF_DEG_Koichi, c(1,7,8,2:4), DF_TF_Koichi, c(1,4,5))

write.table(Features_REL_R$features, "Results/Tables/Genes_of_interest_REL_R.tsv", sep = "\t", row.names = F, quote = F)

Features_REL_R$network %>% 
  igraph::as_data_frame() %>% 
  write.table("Results/Tables/filtered_PPI_TF_REL_R_network.tsv", sep = "\t", row.names = F, quote = F)

Features_REL_R$most_important_network$network %>% 
  igraph::as_data_frame() %>% 
  write.table("Results/Tables/most_important_REL_R_network.tsv", sep = "\t", row.names = F, quote = F)

```

# TCGA

```{r, include = F}
DF_DEG_TCGA <- read.csv("../../REVISION_VD_IDH_Alexis/Results/TCGA_IDHm_DEG.tsv", sep = "\t")
DF_TF_TCGA <- read.csv("../../REVISION_VD_IDH_Alexis/Results/TCGA_IDHm_MS_Viper.tsv", sep = "\t")

Features_TCGA_IDHm_wt <- All_workflow(DF_DEG_TCGA, c(7,1,4), DF_TF_TCGA, c(1,3,4))

write.table(Features_TCGA_IDHm_wt$features, "Results/Tables/Genes_of_interest_TCGA_features.tsv", sep = "\t", row.names = F, quote = F)
Features_TCGA_IDHm_wt$network %>% igraph::as_data_frame() %>% write.table("Results/Tables/filtered_PPI_TF_TCGA_IDHm_network.tsv", sep = "\t", row.names = F, quote = F)

Features_TCGA_IDHm_wt$most_important_network$network %>% 
  igraph::as_data_frame() %>% 
  write.table("Results/Tables/most_important_TCGA_network.tsv", sep = "\t", row.names = F, quote = F)
```

# Verhaak

```{r, include = F}
DF_DEG_Verhaak <- read.csv("../../REVISION_VD_IDH_Alexis/Results/Genes_diff_Verhaak_IDH.csv", sep = "\t")
DF_TF_Verhaak <- read.csv("../../REVISION_VD_IDH_Alexis/Results/Verhaak_IDH_MSVIPER.tsv", sep = "\t")

Features_Verhaak_IDHm_wt <- All_workflow(DF_DEG_Verhaak, c(8, 1, 4), DF_TF_Verhaak, c(1,3,4), logFC_treshold = 0.1)

write.table(Features_Verhaak_IDHm_wt$features, "Results/Tables/Genes_of_interest_Verhaak_features.tsv", sep = "\t", row.names = F, quote = F)
Features_Verhaak_IDHm_wt$network %>% igraph::as_data_frame() %>% write.table("Results/Tables/filtered_PPI_TF_Verhaak_IDHm_network.tsv", sep = "\t", row.names = F, quote = F)

Features_Verhaak_IDHm_wt$most_important_network$network %>% 
  igraph::as_data_frame() %>% 
  write.table("Results/Tables/most_important_Verhaak_network.tsv", sep = "\t", row.names = F, quote = F)
```


```{r, warning=F}
attach(Features_TCGA_IDHm_wt[["features"]])
name <- "TCGA"

base::plot(log(Page_rank), log(Eigen_centrality),
           xlab = "Pagerank (log)",
           main = name,
           ylab = "Eigencentrality (log)")
text(log(Page_rank), log(Eigen_centrality), labels = Gene)

detach(Features_TCGA_IDHm_wt[["features"]])
```

```{r, warning=F}
attach(Features_NR_R$features)
name <- "NR vs R"

base::plot(log(Page_rank), log(Eigen_centrality),
           xlab = "Pagerank (log)",
           main = name,
           ylab = "Eigencentrality (log)")
text(log(Page_rank), log(Eigen_centrality), labels = Gene)

detach(Features_NR_R$features)
```

```{r, warning=F}
attach(Features_REL_R$features)

name <- "REL vs R"

base::plot(log(Page_rank), log(Eigen_centrality),
           xlab = "Pagerank (log)",
           main = name,
           ylab = "Eigencentrality (log)")
text(log(Page_rank), log(Eigen_centrality), labels = Gene)

detach(Features_REL_R$features)
```

```{r, warning=F}
attach(Features_Verhaak_IDHm_wt$features)

name <- "Verhaak"

base::plot(log(Page_rank), log(Eigen_centrality),
           xlab = "Pagerank (log)",
           main = name,
           ylab = "Eigencentrality (log)")
text(log(Page_rank), log(Eigen_centrality), labels = Gene)

detach(Features_Verhaak_IDHm_wt$features)
```

```{r}

```

```{r}
Find_most_importants_genes(Features_NR_R)
```


